Embracing the Laws of Physics: \\ Reversible Models of Computation

### 2017-11-24 survey/survey.tex
keep a bit more. Start on the new story.  Some of the material past the wavefront can be used too, but I have not gotten to it yet.

Our logic of programs and our hardware are based on Boolean Logic
-- which really
ought to be called \emph{Piercean logic}, as Boole had a rather different
conception of logic which is rather far from our current models.
Nevertheless, going back
to Boole's 1853 book entitled \emph{An Investigation of the Laws of Thought,
  on which are Founded the Mathematical Theories of Logic and
  Probabilities}. The opening sentence of Ch.~1 is:
\begin{quote}
  The design of the following treatise is to investigate the fundamental laws
  of those operations of the mind by which reasoning is performed; \ldots
\end{quote}
A few chapters later, we find:
\begin{quote}
  \textbf{Proposition IV.}  That axiom of metaphysicians which is termed the
  principle of contradiction, and which affirms that it is impossible for any
  being to possess a quality, and at the same time not to possess it, is a
  consequence of the fundamental law of thought, whose expression is $x^2 =
  x$.
\end{quote}
A detailed historical analysis of Boole's ideas are beyond our scope.
The above quotes, however, should convey the idea that our notions of
computation generally date back to ideas that were thought reasonable
before \tilde 1900.

@@

A physical principle of computation

No creation of information
No duplication of information

@@

When a physical law equates two quantities like force and the product of mass
and acceleration, a deep understanding of the law demands that we interpret
the equality as an isomorphism, i.e., as two processes, one in each
direction, that transform one side of the equation to the other. In fact, the
same insight applies to constructive mathematics and its homotopy
interpretation. There is a constructive character to equality that is at the
heart of the constructive approach to mathematics.

>>>>

Our logic of programs and our hardware are based on Boolean Logic
-- which really
ought to be called \emph{Piercean logic}, as Boole had a rather different
conception of logic which is rather far from our current models.
Nevertheless, going back
to Boole's 1853 book entitled \emph{An Investigation of the Laws of Thought,
  on which are Founded the Mathematical Theories of Logic and
  Probabilities}. The opening sentence of Ch.~1 is:
\begin{quote}
  The design of the following treatise is to investigate the fundamental laws
  of those operations of the mind by which reasoning is performed; \ldots
\end{quote}
A few chapters later, we find:
\begin{quote}
  \textbf{Proposition IV.}  That axiom of metaphysicians which is termed the
  principle of contradiction, and which affirms that it is impossible for any
  being to possess a quality, and at the same time not to possess it, is a
  consequence of the fundamental law of thought, whose expression is $x^2 =
  x$.
\end{quote}
A detailed historical analysis of Boole's ideas are beyond our scope.
The above quotes, however, should convey the idea that our notions of
computation generally date back to ideas that were thought reasonable
before \textasciitilde 1900.

@@

A high level approach to conservation of information would proceed
as follows:
\begin{enumerate}
\item Create a means to measure information.
\item Show that this measure is sound with respect to our current understanding
of what information ought to be.
\item Assert that whatever ``quantity'' of information a system has, must
remain invariant throughout the lifetime of the system.
\end{enumerate}

Certainly, whatever our notion of information, it can neither be created,
duplicated, nor erased. Of course, if we make our notion of information
too rigid, it will be impossible to even \textbf{modify} it! For computation
to occur, modification must be possible.  But of what kind?

\jc{wavefront}

Example : $2+5 = 5+2$. Green and Red apples versus apples. What
can we observe? If we can't observe something, then we call them
the same (Leibniz).

@@
